\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{longtable}
\usepackage{array}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{booktabs}
\usepackage{listings}

\definecolor{primary}{HTML}{5B3E7A}
\definecolor{muted}{HTML}{666666}
\definecolor{codebg}{HTML}{F7F7F8}

\hypersetup{
  colorlinks=true,
  linkcolor=primary,
  urlcolor=primary
}

\pagestyle{fancy}
\fancyhf{}
\lhead{E-Commerce Backend Blueprint}
\rhead{Node.js + Express + MongoDB}
\cfoot{\thepage}

\titleformat{\section}{\large\bfseries\color{primary}}{\thesection}{0.6em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.6em}{}
\setlist[itemize]{leftmargin=1.3em}
\setlist[enumerate]{leftmargin=1.5em}

\lstdefinestyle{code}{
  backgroundcolor=\color{codebg},
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
  rulecolor=\color{codebg},
  columns=fullflexible
}

\title{\textbf{Scalable Backend Plan for Stylee E-Commerce}}
\author{Prepared for Node.js Backend Implementation}
\date{May 21, 2026}

\begin{document}

\maketitle
\tableofcontents
\newpage

\section{Project Summary}

Your frontend currently has the following user-facing flows:

\begin{itemize}
  \item Home page, shop page, category filtering, product search, product details, related products.
  \item User login, register, profile update, saved addresses, and order history.
  \item Cart, quantity update, promo code, wishlist, checkout, and contact form.
  \item Current data source is mostly static JavaScript files and browser localStorage.
\end{itemize}

To make this project scalable, the backend should move these responsibilities to secure APIs, a database, file storage, payment gateway integrations, and admin controls.

\section{How Many APIs Are Needed?}

\subsection{Recommended Endpoint Count}

\begin{longtable}{p{0.25\linewidth} p{0.18\linewidth} p{0.48\linewidth}}
\toprule
\textbf{Stage} & \textbf{Approx. APIs} & \textbf{Purpose} \\
\midrule
MVP & 35--45 & Enough for user auth, products, categories, cart, wishlist, checkout, orders, address book, and contact form. \\
Scalable v1 & 60--75 & Adds admin product management, coupons, reviews, payment webhooks, inventory control, uploads, analytics, and order status management. \\
Advanced v2 & 90+ & Adds returns, refunds, notifications, recommendations, vendor management, audit logs, and warehouse flows. \\
\bottomrule
\end{longtable}

\textbf{Recommendation:} Start with the MVP plus a clean scalable structure. Do not build every advanced feature on day one, but design the folders, models, and middleware so new modules can be added without rewriting the backend.

\section{Backend Technology Stack}

\subsection{Core Stack}

\begin{longtable}{p{0.28\linewidth} p{0.62\linewidth}}
\toprule
\textbf{Technology} & \textbf{Use} \\
\midrule
Node.js & JavaScript runtime for the backend. \\
Express.js & REST API framework. \\
MongoDB & Main database for products, users, orders, carts, and wishlist. \\
Mongoose & MongoDB schema modeling and validation. \\
JWT & Access-token based authentication. \\
HTTP-only cookies & Safer token storage for web applications. \\
Cloudinary or S3 & Product image and avatar uploads. \\
Stripe or Razorpay & Payment processing. Razorpay is usually better for India-first stores. \\
Nodemailer or Resend & Transactional emails such as order confirmation and contact replies. \\
\bottomrule
\end{longtable}

\subsection{Important Note About Current Backend}

Your current \texttt{e-backend/package.json} includes \texttt{fs}. That package should be removed because Node.js already provides \texttt{fs} as a built-in module. You do not need to install it from npm.

\section{Recommended Dependencies}

\subsection{Production Dependencies}

\begin{lstlisting}[style=code]
npm install express mongoose dotenv cors helmet cookie-parser jsonwebtoken bcryptjs zod express-rate-limit morgan compression multer cloudinary nodemailer slugify nanoid
\end{lstlisting}

\subsection{Payment Dependency}

Choose one payment gateway first:

\begin{lstlisting}[style=code]
# Option 1: India-first payments
npm install razorpay

# Option 2: International card payments
npm install stripe
\end{lstlisting}

\subsection{Development Dependencies}

\begin{lstlisting}[style=code]
npm install -D nodemon eslint prettier jest supertest
\end{lstlisting}

\subsection{Dependency Purpose}

\begin{longtable}{p{0.28\linewidth} p{0.62\linewidth}}
\toprule
\textbf{Package} & \textbf{Purpose} \\
\midrule
express & API server and routing. \\
mongoose & MongoDB models and queries. \\
dotenv & Environment variables. \\
cors & Allow frontend to call backend safely. \\
helmet & Security headers. \\
cookie-parser & Read HTTP-only auth cookies. \\
jsonwebtoken & Generate and verify auth tokens. \\
bcryptjs & Hash user passwords. \\
zod & Request body validation. \\
express-rate-limit & Protect login/contact APIs from spam. \\
morgan & Request logging in development. \\
compression & Gzip/Brotli-style response compression. \\
multer & Receive uploaded files. \\
cloudinary & Store product/avatar images. \\
nodemailer & Send emails. \\
slugify & Create SEO-friendly product/category slugs. \\
nanoid & Generate short public IDs where needed. \\
razorpay/stripe & Payment order creation and webhook verification. \\
nodemon & Restart server during development. \\
jest + supertest & API testing. \\
eslint + prettier & Code quality and formatting. \\
\bottomrule
\end{longtable}

\section{Professional Folder Structure}

\begin{lstlisting}[style=code]
e-backend/
  src/
    app.js
    server.js
    config/
      db.js
      env.js
      cloudinary.js
      payment.js
    modules/
      auth/
        auth.controller.js
        auth.routes.js
        auth.service.js
        auth.validation.js
      users/
        user.model.js
        user.controller.js
        user.routes.js
        user.service.js
      addresses/
        address.model.js
        address.controller.js
        address.routes.js
      categories/
        category.model.js
        category.controller.js
        category.routes.js
      products/
        product.model.js
        product.controller.js
        product.routes.js
        product.validation.js
      cart/
        cart.model.js
        cart.controller.js
        cart.routes.js
      wishlist/
        wishlist.model.js
        wishlist.controller.js
        wishlist.routes.js
      coupons/
        coupon.model.js
        coupon.controller.js
        coupon.routes.js
      orders/
        order.model.js
        order.controller.js
        order.routes.js
      payments/
        payment.controller.js
        payment.routes.js
        payment.service.js
      reviews/
        review.model.js
        review.controller.js
        review.routes.js
      contact/
        contact.model.js
        contact.controller.js
        contact.routes.js
      admin/
        admin.routes.js
        admin.controller.js
    middleware/
      auth.middleware.js
      admin.middleware.js
      error.middleware.js
      validate.middleware.js
      upload.middleware.js
      rateLimit.middleware.js
    utils/
      ApiError.js
      ApiResponse.js
      asyncHandler.js
      generateToken.js
      sendEmail.js
      pagination.js
    seeds/
      seedProducts.js
      seedCategories.js
    tests/
      auth.test.js
      products.test.js
      orders.test.js
  .env
  .env.example
  package.json
\end{lstlisting}

\section{Database Models}

\subsection{User Model}

\begin{itemize}
  \item \texttt{name}: String, required.
  \item \texttt{email}: String, required, unique, lowercase.
  \item \texttt{password}: String, required, hashed.
  \item \texttt{phone}: String.
  \item \texttt{avatar}: String.
  \item \texttt{role}: \texttt{user} or \texttt{admin}.
  \item \texttt{isEmailVerified}: Boolean.
  \item \texttt{createdAt}, \texttt{updatedAt}.
\end{itemize}

\subsection{Product Model}

\begin{itemize}
  \item \texttt{name}, \texttt{slug}, \texttt{description}.
  \item \texttt{category}: ObjectId reference to Category.
  \item \texttt{gender}: Women, Men, Unisex.
  \item \texttt{price}, \texttt{discountPrice}.
  \item \texttt{images}: Array of image URLs.
  \item \texttt{sizes}: Array.
  \item \texttt{colors}: Array.
  \item \texttt{stock}: Number.
  \item \texttt{ratingAverage}, \texttt{ratingCount}.
  \item \texttt{isActive}: Boolean.
\end{itemize}

\subsection{Category Model}

\begin{itemize}
  \item \texttt{name}: String.
  \item \texttt{slug}: String, unique.
  \item \texttt{image}: String.
  \item \texttt{description}: String.
  \item \texttt{isActive}: Boolean.
\end{itemize}

\subsection{Cart Model}

\begin{itemize}
  \item \texttt{user}: ObjectId reference to User.
  \item \texttt{items}: product, size, color, quantity, priceSnapshot.
  \item \texttt{coupon}: optional coupon code.
  \item \texttt{subtotal}, \texttt{shipping}, \texttt{discount}, \texttt{total}.
\end{itemize}

\subsection{Wishlist Model}

\begin{itemize}
  \item \texttt{user}: ObjectId reference to User.
  \item \texttt{products}: Array of Product ObjectIds.
\end{itemize}

\subsection{Order Model}

\begin{itemize}
  \item \texttt{orderNumber}: Public order ID such as \texttt{ORD-98421}.
  \item \texttt{user}: ObjectId reference to User.
  \item \texttt{items}: product, name, image, size, color, quantity, price.
  \item \texttt{shippingAddress}: embedded address snapshot.
  \item \texttt{payment}: gateway, paymentId, orderId, status.
  \item \texttt{pricing}: subtotal, shipping, discount, tax, total.
  \item \texttt{status}: Pending, Processing, Shipped, Delivered, Cancelled.
  \item \texttt{deliveredAt}, \texttt{cancelledAt}.
\end{itemize}

\section{REST API Blueprint}

\subsection{Health}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
GET & /api/health & Check backend status. \\
\bottomrule
\end{longtable}

\subsection{Auth APIs}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
POST & /api/auth/register & Create a new user account. \\
POST & /api/auth/login & Login and set auth cookie/token. \\
POST & /api/auth/logout & Logout and clear cookie/token. \\
GET & /api/auth/me & Get current logged-in user. \\
POST & /api/auth/refresh & Refresh access token. \\
POST & /api/auth/forgot-password & Send reset password email. \\
POST & /api/auth/reset-password & Reset password using token. \\
\bottomrule
\end{longtable}

\subsection{User and Profile APIs}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
GET & /api/users/profile & Get user profile. \\
PATCH & /api/users/profile & Update name, phone, avatar. \\
PATCH & /api/users/password & Change password. \\
\bottomrule
\end{longtable}

\subsection{Address APIs}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
GET & /api/addresses & Get logged-in user's addresses. \\
POST & /api/addresses & Add new address. \\
PATCH & /api/addresses/:id & Update address. \\
PATCH & /api/addresses/:id/default & Set default address. \\
DELETE & /api/addresses/:id & Delete address. \\
\bottomrule
\end{longtable}

\subsection{Category APIs}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
GET & /api/categories & Public category list. \\
GET & /api/categories/:slug & Category detail. \\
POST & /api/admin/categories & Admin create category. \\
PATCH & /api/admin/categories/:id & Admin update category. \\
DELETE & /api/admin/categories/:id & Admin delete/deactivate category. \\
\bottomrule
\end{longtable}

\subsection{Product APIs}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
GET & /api/products & Product list with search, filters, sort, pagination. \\
GET & /api/products/:slug & Product detail. \\
GET & /api/products/:id/related & Related products. \\
POST & /api/admin/products & Admin create product. \\
PATCH & /api/admin/products/:id & Admin update product. \\
PATCH & /api/admin/products/:id/stock & Admin update stock. \\
DELETE & /api/admin/products/:id & Admin delete/deactivate product. \\
\bottomrule
\end{longtable}

\subsection{Product Query Parameters}

\begin{lstlisting}[style=code]
GET /api/products?category=Dresses&gender=Women&size=M&color=Lavender&minPrice=1000&maxPrice=15000&search=silk&sort=price-low&page=1&limit=12
\end{lstlisting}

\section{Cart APIs}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
GET & /api/cart & Get user's cart. \\
POST & /api/cart/items & Add item to cart. \\
PATCH & /api/cart/items/:cartItemId & Update item quantity. \\
DELETE & /api/cart/items/:cartItemId & Remove item from cart. \\
DELETE & /api/cart & Clear cart. \\
POST & /api/cart/apply-coupon & Apply promo code. \\
DELETE & /api/cart/coupon & Remove promo code. \\
\bottomrule
\end{longtable}

\section{Wishlist APIs}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
GET & /api/wishlist & Get wishlist. \\
POST & /api/wishlist/:productId & Add product to wishlist. \\
DELETE & /api/wishlist/:productId & Remove product from wishlist. \\
DELETE & /api/wishlist & Clear wishlist. \\
\bottomrule
\end{longtable}

\section{Coupon APIs}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
POST & /api/coupons/validate & Validate coupon for cart. \\
GET & /api/admin/coupons & Admin list coupons. \\
POST & /api/admin/coupons & Admin create coupon. \\
PATCH & /api/admin/coupons/:id & Admin update coupon. \\
DELETE & /api/admin/coupons/:id & Admin deactivate coupon. \\
\bottomrule
\end{longtable}

\section{Checkout, Payment, and Order APIs}

\subsection{Checkout and Payments}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
POST & /api/checkout/summary & Calculate final cart summary. \\
POST & /api/payments/create-order & Create Razorpay/Stripe payment order. \\
POST & /api/payments/verify & Verify client payment result. \\
POST & /api/payments/webhook & Gateway webhook for payment status. \\
\bottomrule
\end{longtable}

\subsection{Orders}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
POST & /api/orders & Create order after payment success or COD selection. \\
GET & /api/orders/my & User order history. \\
GET & /api/orders/:id & User order detail. \\
PATCH & /api/orders/:id/cancel & User cancel order if allowed. \\
GET & /api/admin/orders & Admin list all orders. \\
PATCH & /api/admin/orders/:id/status & Admin update status. \\
\bottomrule
\end{longtable}

\section{Review APIs}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
GET & /api/products/:productId/reviews & Get product reviews. \\
POST & /api/products/:productId/reviews & Add review after purchase. \\
PATCH & /api/reviews/:id & Update own review. \\
DELETE & /api/reviews/:id & Delete own review or admin delete. \\
\bottomrule
\end{longtable}

\section{Contact and Support APIs}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
POST & /api/contact & Submit contact form. \\
GET & /api/admin/contact & Admin list inquiries. \\
PATCH & /api/admin/contact/:id/status & Admin update inquiry status. \\
\bottomrule
\end{longtable}

\section{Upload APIs}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
POST & /api/uploads/image & Upload product/category/avatar image. \\
DELETE & /api/uploads/image & Delete image from storage. \\
\bottomrule
\end{longtable}

\section{Admin Dashboard APIs}

\begin{longtable}{p{0.16\linewidth} p{0.34\linewidth} p{0.40\linewidth}}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Purpose} \\
\midrule
GET & /api/admin/stats & Sales, users, orders, revenue summary. \\
GET & /api/admin/users & Admin list users. \\
PATCH & /api/admin/users/:id/role & Change user role. \\
PATCH & /api/admin/users/:id/status & Block/unblock user. \\
\bottomrule
\end{longtable}

\section{Environment Variables}

\begin{lstlisting}[style=code]
NODE_ENV=development
PORT=5000
CLIENT_URL=http://localhost:5173

MONGO_URI=mongodb://127.0.0.1:27017/stylee_ecommerce

JWT_ACCESS_SECRET=replace_with_strong_access_secret
JWT_REFRESH_SECRET=replace_with_strong_refresh_secret
JWT_ACCESS_EXPIRES=15m
JWT_REFRESH_EXPIRES=7d

CLOUDINARY_CLOUD_NAME=
CLOUDINARY_API_KEY=
CLOUDINARY_API_SECRET=

RAZORPAY_KEY_ID=
RAZORPAY_KEY_SECRET=
RAZORPAY_WEBHOOK_SECRET=

SMTP_HOST=
SMTP_PORT=
SMTP_USER=
SMTP_PASS=
EMAIL_FROM=Stylee <support@stylee.com>
\end{lstlisting}

\section{Standard API Response Format}

\begin{lstlisting}[style=code]
{
  "success": true,
  "message": "Product fetched successfully",
  "data": {},
  "meta": {
    "page": 1,
    "limit": 12,
    "total": 100
  }
}
\end{lstlisting}

\section{Security Checklist}

\begin{itemize}
  \item Hash passwords using \texttt{bcryptjs}.
  \item Store tokens in HTTP-only cookies where possible.
  \item Use \texttt{helmet} for security headers.
  \item Use \texttt{cors} with only your frontend domain.
  \item Use rate limiting on login, register, contact, and forgot-password APIs.
  \item Validate every request body using \texttt{zod}.
  \item Never trust product price from frontend during checkout.
  \item Recalculate subtotal, discount, shipping, and total on backend.
  \item Reduce stock only after payment verification.
  \item Verify payment webhooks using gateway signature.
  \item Add admin middleware for every \texttt{/api/admin/*} route.
\end{itemize}

\section{Frontend Replacement Plan}

\begin{longtable}{p{0.32\linewidth} p{0.58\linewidth}}
\toprule
\textbf{Current Frontend Source} & \textbf{Backend Replacement} \\
\midrule
\texttt{src/data/products.js} & \texttt{GET /api/products} and \texttt{GET /api/products/:slug}. \\
\texttt{src/data/categories.js} & \texttt{GET /api/categories}. \\
\texttt{AuthContext localStorage} & \texttt{/api/auth/*}, \texttt{/api/users/profile}, \texttt{/api/addresses}. \\
\texttt{CartContext localStorage} & \texttt{/api/cart/*}. \\
\texttt{WishlistContext localStorage} & \texttt{/api/wishlist/*}. \\
\texttt{placeOrder mock function} & \texttt{/api/payments/*} and \texttt{/api/orders}. \\
\texttt{Contact setTimeout mock} & \texttt{POST /api/contact}. \\
\bottomrule
\end{longtable}

\section{Implementation Order}

\begin{enumerate}
  \item Create backend folder structure and common utilities.
  \item Connect MongoDB using Mongoose.
  \item Add global error handler, async handler, and validation middleware.
  \item Build auth: register, login, logout, current user.
  \item Build category and product APIs.
  \item Seed current frontend products and categories into MongoDB.
  \item Build cart and wishlist APIs.
  \item Build address APIs and profile APIs.
  \item Build coupon validation and backend pricing calculation.
  \item Build payment creation, payment verification, and order creation.
  \item Build admin product/order/category controls.
  \item Connect frontend contexts to backend APIs.
  \item Add tests for auth, products, cart, and orders.
  \item Deploy backend and configure production environment variables.
\end{enumerate}

\section{MVP Route Count Summary}

\begin{longtable}{p{0.40\linewidth} p{0.20\linewidth} p{0.30\linewidth}}
\toprule
\textbf{Module} & \textbf{MVP APIs} & \textbf{Priority} \\
\midrule
Health & 1 & High \\
Auth & 5--7 & High \\
User/Profile & 2--3 & High \\
Addresses & 5 & High \\
Categories & 2 & High \\
Products & 3 & High \\
Cart & 7 & High \\
Wishlist & 4 & Medium \\
Coupons & 1--2 & Medium \\
Checkout/Payments & 3--4 & High \\
Orders & 4 & High \\
Contact & 1 & Medium \\
Admin basic & 8--12 & Medium \\
\bottomrule
\end{longtable}

\section{Final Recommendation}

For your current frontend, build around \textbf{45 MVP APIs} first. Keep the folder structure ready for around \textbf{70 scalable APIs}. The most important backend rule is this: cart totals, coupon discount, stock validation, payment verification, and order creation must happen on the backend, not only in React.

\end{document}
